Model Recognition as Planning

### 2018-09-20 doc/icaps-19/model-recognition.tex
paper icaps19: pages 1,2 and 3 stable. Ready for review

\section{Recognition of \strips\ models}
\label{sec:asPlanning}
Here we analyze the particular instantiation of the {\em model recognition} task where the semantics of the actions $A$ (i.e. the $\rho$ and $\theta$ functions) are specified with \strips\ action schemas. We first formalize \strips\ schemas, then define the full space of possible \strips\ schema and eventually, we introduce an {\em edit distance} to estimate the $P(\mathcal{M}|\tau)$ probabilities for \strips\ models.

@@

\subsection{A propositional encoding for \strips\ action schema}
\strips\ action schema provide a compact representation for specifying classical planning action models. In more detail, a \strips\ action schema $\xi$ with name, $name(\xi)$, defines a list of {\em parameters} $pars(\xi)$, and the three list of predicates ($pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schema. Figure~\ref{fig:model-example} shows six action schema codded in PDDL for a robot navigation in a $n\times n$ grid.

We say that two \strips\ schemes $\xi$ and $\xi'$ are {\em comparable} iff both share the same list of parameters. For instance, we claim that the six action schema of Figure~\ref{fig:model-example} are {\em comparable} while {\small\tt stack(?v1,?v2)} and {\small\tt pickup(?v1)} from a four opertator {\em blocksworld}~\cite{slaney2001blocks} are not. Last but not least, two \strips\ models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every action schema $\xi\in\mathcal{M}$ to a comparable schema $\xi'\in\mathcal{M'}$ and vice versa.

@@

\subsection{Experimental setup}
We randomly generated a $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_{100}\}$ set of one-hundred different {\em Turing Machines} where each $\mathcal{M}\in M$ is a seven-symbol six-state {\em Turing Machine}. The classical planning frame $\Phi=\tup{F,A}$ is the same for all the {\em Turing Machines}, there is an $a\in A$ action for each pair of tape symbol and non-terminal state machine, while the $\theta(s,a)$ function is defined differently for each the machines (using a different \strips\ action model).

@@

>>>>

\section{Recognition of \strips\ models}
\label{sec:asPlanning}
Here we analyze the particular instantiation of the {\em model recognition} task where the semantics of actions (in other words, the corresponding $\rho$ and $\theta$ functions) are specified with \strips\ action schemas.

We first formalize \strips\ schemas, then define the full space of possible \strips\ schema and eventually, we introduce an {\em edit distance} which allows us to estimate the $P(\tau|\mathcal{M})$ likelihoods for observations of the execution of classical plans and planning \strips\ models.

@@

\subsection{A propositional encoding for \strips\ action schema}
\strips\ action schema provide a compact representation for specifying classical planning models. For instance, Figure~\ref{fig:model-example} shows six \strips\ action schema codded in PDDL for a robot navigation in a $n\times n$ grid. A \strips\ action schema $\xi$ with name, $name(\xi)$, defines a list of {\em parameters} $pars(\xi)$, and the three list of predicates ($pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schema.

We say that two \strips\ schemes $\xi$ and $\xi'$ are {\em comparable} iff $pars(\xi)=pars(\xi')$, both share the same list of parameters. For instance, we claim that the six action schema of Figure~\ref{fig:model-example} are {\em comparable} while the {\small\tt stack(?v1,?v2)} and {\small\tt pickup(?v1)} schemes from the four opertator {\em blocksworld}~\cite{slaney2001blocks} are not. Last but not least, we say that two \strips\ models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every action schema $\xi\in\mathcal{M}$ to a comparable schema $\xi'\in\mathcal{M'}$ and vice versa.

@@

\subsection{Experimental setup}
We randomly generated a $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_{100}\}$ set of one-hundred different {\em Turing Machines} where each $\mathcal{M}\in M$ is a seven-symbol six-state {\em Turing Machine}. The classical planning frame $\tup{F,A}$ is the same for all the {\em Turing Machines}, there is an $a\in A$ action for each pair of tape symbol and non-terminal state machine, while the $\rho$ and $\theta$ function is defined differently for each the machines (using a different \strips\ action model).

@@

Remarkably, the extension of this piece of work to the FOND planning setting~\cite{muise2012improved} is straightforward by simply considering the {\em all-outcomes} determiniztion of the actions with non-determinitic effects~\cite{yoon2007ff}. An interesting research direction is however to understand how to apply our approach to planning models where the planning models include actions with probabilistic effects~\cite{younes2005first}.
